Enhancing Adversarial Robustness for Deep Metric Learning

### 2021-10-26 robdml.tex
tex

To this end, we propose Graduate Adversary, a family of pseudo-hardness
functions which can be used as the \emph{destination} hardness in HM.
%
Any function that leads to a relatively weak adversary in the early training
phase, and a relatively strong adversary in the late training phase belongs to
this family.
%
As an example, we design a ``Simple Graduate Adversary'' (SGA) pseudo-hardness
function as the linearly scaled negative triplet margin, incorporating a strong
assumption that the \emph{destination} hardness should remain Semihard.
%
%Empirical observation will show the advantage of such pseudo-hardness
%functions.

@@

% contributions
In brief, our contribusions includes:
%
\begin{enumerate}[nosep, noitemsep, leftmargin=*]
	%
	\item {\textit{Hardness Manipulation}} (HM) is proposed as a flexible tool
		to create adversarial triplets with increased hardness for
		subsequent adversarial training.
		%
		It is more efficient than previous adversarial training methods of DML.
		%
	\item \textit{Gradual Adversary} is proposed as a family of pseudo-hardness
		functions for \emph{Hardness Manipulation}, which can balance the
		training objectives during the training process.
		%
		A linear function named Simple Graduate Adversary (SGA) is designed as
		an example.
		%
	\item \textit{Intra-Class Structure} penalty is proposed to further improve
		the robustness of a DML model, while such possibility is neglected by
		existing methods. 
		%
\end{enumerate}

@@

Thus, a desired $H_\mathsf{D}$ should be (1) close to $H_\mathsf{S}$ in the
early training phase (indicated by a large loss value), and (2) clearly higher
than $H_\mathsf{S}$ in the late training phase (indicated by a small loss
value).
%
We name the two conditions as ``Graduate Adversary'' for 
pseudo-hardness functions.

Additionally, if we have a strong prior knowledge on ``how hard the destination
hardness should be'', the pseudo-hardness function can be independent from any
benign triplet.

% simple graduate adversary

For instance, if we assume that the destination hardness should remain
in the hardness range of Semihard~\cite{facenet}, the following function
satisfies the two condictions of Gradual Adversary, dubbed
``Simple Gradual Adversary'' (SGA):
%
\begin{equation}
	%
	g_\mathsf{SGA}(\bar{\ell}_{t-1}) =
	-\gamma \cdot \bar{\ell}_{t-1} ~ \in [-\gamma,0].
	%
\end{equation}
%
And the adversarial examples from $\text{HM}(H_\mathsf{S},g_\mathsf{SGA})$
can be used for adversarial training.
%
The design of this function will be empirically supported in the experiments
section.
%
\cref{fig:ga}
%
The exploration of other possible pseudo-hardness functions is omitted 
for simplicity.

@@

During adversarial training, the adversarial counterpart of each given sample
triplet is fed to the model, and the triplet loss will enforce a good
\emph{inter-class} structure.
%
Since the anchor, positive sample, and their adversarial counterpart belongs to
the same class, it should be noted that the \emph{intra-class} structure can be
enforced as well, and This has been neglected by the existing DML defenses.
%
\emph{Intra-class} structure is also important for robustness besides the
\emph{inter-class} structure, because the attack may attempt to change the
rankings of samples in the same class~\cite{advrank}.

>>>>

To this end, we propose Graduate Adversary, a family of pseudo-hardness
functions which can be used as the \emph{destination} hardness in HM.
%
Any function that leads to a relatively weak adversary in the early training
phase, and a relatively strong adversary in the late training phase belongs to
this family.
%
As an example, we design a ``Linear Graduate Adversary'' (LGA) pseudo-hardness
function as the linearly scaled negative triplet margin, incorporating a strong
assumption that the \emph{destination} hardness should remain Semihard.
%
%Empirical observation will show the advantage of such pseudo-hardness
%functions.

@@

% contributions
In brief, our contribusions includes:
%
\begin{enumerate}[nosep, noitemsep, leftmargin=*]
	%
	\item {\textit{Hardness Manipulation}} (HM) is proposed as a flexible tool
		to create adversarial triplets with increased hardness for
		subsequent adversarial training.
		%
		It is more efficient than previous adversarial training methods of DML.
		%
	\item \textit{Gradual Adversary} is proposed as a family of pseudo-hardness
		functions for \emph{Hardness Manipulation}, which can balance the
		training objectives during the training process.
		%
		A linear function named Linear Graduate Adversary (LGA) is designed as
		an example.
		%
	\item \textit{Intra-Class Structure} penalty is proposed to further improve
		the robustness of a DML model, while such possibility is neglected by
		existing methods. 
		%
\end{enumerate}

@@

Thus, $H_\mathsf{D}$ should be (1) close to $H_\mathsf{S}$ in the
early training phase (indicated by a large loss value), and (2) clearly higher
than $H_\mathsf{S}$ in the late training phase (indicated by a small loss
value),
%
in order to automatically balance the training objectives (\ie, performance
on benign examples \emph{v.s.} robustness).
%
We name the two conditions as ``Graduate Adversary'' for 
pseudo-hardness functions, as illustrated in \cref{fig:ga}.

% Linear graduate adversary

As an example, we propose a ``Linear Gradual Adversary'' (LGA) pseudo-hardness
function that is independent to any benign triplets, incorporating an
assumption that $H_\mathsf{D}$ should remain Semihard~\cite{facenet}, as
follows:
%
\begin{equation}
	%
	g_\mathsf{LGA}(\bar{\ell}_{t-1}) = -\gamma \cdot \bar{\ell}_{t-1} ~ \in
	[-\gamma,0].
	%
\end{equation}
%
And the training objectives will be automatically balanced in
$\text{HM}[H_\mathsf{S},g_\mathsf{LGA}]$.
%
The design choice of LGA is empirically supported by experimental observations.
%
We leave more complex pseudo-hardness functions for future study.
%

@@

During adversarial training with HM, the adversarial counterpart of each given sample
triplet is fed to the model, and the triplet loss will enforce a good
\emph{inter-class} structure.
%
Since the anchor, positive sample, and their adversarial counterpart belongs to
the same class, it should be noted that the \emph{intra-class} structure can be
enforced as well, and This has been neglected by the existing DML defenses.
%
\emph{Intra-class} structure is also important for robustness besides the
\emph{inter-class} structure, because the attack may attempt to change the
rankings of samples in the same class~\cite{advrank}.
